Pillar 1 - Conformal Prediction

### 2019-05-22 Pillar 1 - Conformal Prediction.Rmd
update parameters of plots for COPA2019 editting

@@

```{r}
course_selection <- c("COR1002", "COR1004", "SCI3003", "SCI2040", "SSC3044", "SSC3038", "SCI2018", "SCI2010")

fit_lasso %>% 
  filter(target %in% course_selection) %>% 
  select(target, n, cv_error, df)
```

@@

create_training_temp <- function(df, fold_current) df %>% filter(fold != fold_current)

@@

  n <- nrow(df)
  n_training      <- floor(0.67 * n) # 2/3
  n_callibration  <- n - n_training  # 1/3

  allocation <- rep(x     = c("training", "callibration"), 
                    times = c(n_training, n_callibration)) %>%
  sample

@@

create_training     <- function(df) df %>% filter(allocation == "training")

create_callibration <- function(df) df %>% filter(allocation == "callibration")
```

@@

  cv.glmnet(x = as.matrix(X), y = Y, type.measure = "mae") # lasso by default

@@

  Y_hat <- predict.cv.glmnet(model, newx = as.matrix(X), s = "lambda.min")

  df %>% mutate(Y_hat = as.vector(Y_hat),
                error = abs(Y_hat - Y))

@@

  X <- df %>% select(matches("GPA|Topic"))
  error <- df %>% pull(error)

  cv.glmnet(x = as.matrix(X), y = error, type.measure = "mae") # lasso by default

@@

  error_hat <- predict.cv.glmnet(model, newx = as.matrix(X), s = "lambda.min")

  df %>% mutate(error_hat = as.vector(error_hat))

@@

## Compute non-conformity scores on callibration test
```{r}
compute_ncs <- function(df) df %>% mutate(alpha = abs(Y - Y_hat) / error_hat )

compute_confidence <- function(df) df %>% mutate(confidence = rank(alpha) / nrow(.) )
```

## Evaluate coverage on test set

```{r}
compute_width <- function(df, confidence_given){

  df <- df %>% 
    distinct(alpha, confidence) %>%
    filter(confidence >= confidence_given)

@@

  df %>% mutate(
    width       = ncs * error_hat,
    border_low  = Y_hat - width,
    border_high = Y_hat + width
    )
}

@@

```{r, echo = FALSE}
for(course in course_selection){

  # Set up
  callibration <- list()
  coverage     <- list()

  confidence <- seq(0.01, 0.99, 0.01)
  df_coverage <- tibble::tibble(
    confidence = confidence,
    width_mean = NA,
    coverage   = NA
    )

@@

    # training, callibration and test sets
    Z_test          <- create_test(Z, fold_current = iteration)
    Z_training_temp <- create_training_temp(Z, fold_current = iteration) %>% add_allocation
    Z_training      <- create_training(Z_training_temp)
    Z_callibration  <- create_callibration(Z_training_temp)

    # fit target and error model to training set
    model_target <- fit_target_model(Z_training)
    Z_training   <- predict_grade(Z_training, model_target)
    model_error  <- fit_error_model(Z_training)

    # non-conformity scores on callibration
    Z_callibration <- Z_callibration %>% 
      predict_grade(model_target) %>%
      predict_error(model_error) %>%
      compute_ncs %>%
      compute_confidence

    # coverage on test set
    Z_test <- Z_test %>% 
      predict_grade(model_target) %>%
      predict_error(model_error)

    for(i in 1 : nrow(df_coverage)){ # loop through confidence levels

      confidence_i <- df_coverage$confidence[i]
      ncs          <- compute_width(Z_callibration, confidence_i)

      Z_test <- Z_test %>% compute_interval(ncs = ncs) %>% compute_hit

      coverage_i   <- mean(Z_test$hit)
      width_mean_i <- mean(Z_test$width)

      df_coverage$width_mean[i] <- width_mean_i
      df_coverage$coverage[i]   <- coverage_i

    # Output fold
    callibration[[iteration]] <- Z_callibration %>% mutate(iteration = iteration) %>% select(iteration, confidence, alpha)
    coverage[[iteration]]     <- df_coverage %>% mutate(iteration = iteration)

@@

  # Output

  callibration <- bind_rows(callibration) %>% 
    mutate(iteration = as.factor(iteration))
  coverage     <- bind_rows(coverage) %>%
    mutate(iteration = as.factor(iteration))

  title_course <-  str_c(course, " (n = ", nrow(Z), ")")
  subtitle     <- "A line per CV test sets"

  print(ggplot(data = coverage) +
          geom_hline(yintercept = 1, linetype = "dashed") +
          geom_line(aes(confidence, width_mean, group = iteration, col = iteration), size = .5) +
          labs(title = title_course,
               subtitle = subtitle) +
    theme(legend.position = "none"))

  print(ggplot(data = coverage) +
          geom_abline(slope = 1, linetype = "dashed") +
          geom_line(aes(confidence, coverage, group = iteration, col = iteration), size = .5) +
          labs(title = title_course,
               subtitle = subtitle) +
    theme(legend.position = "none"))
}
```

Illustration of the effect of error_hat on interval's width (for confidence of 99%):
```{r, fig.height=4}
Z_test %>% select(Y, Y_hat, error, error_hat, width, border_low, border_high, hit)
plot(Z_test$error, Z_test$error_hat)
```

>>>>

```{r setup_graph}
theme_set(theme_light())

output_folder <- "Conference/COPA2019/paper/figures"
plot_width  <- 6
plot_height <- plot_width / 1.6 
```

@@

```{r course_selection}
course_selection <- c(
#  "SCI3003", "SCI2040", # remove because n too small
  "COR1002", "COR1004",  "SSC3044", "SSC3038", "SCI2018", "SCI2010")

course_table <- fit_lasso %>% 
  filter(target %in% course_selection) %>% 
  select(target, n, cv_error, df)
```

@@

create_training <- function(df, fold_current) df %>% filter(fold != fold_current)

@@

  n_training <- nrow(df)
  n_proper_training <- floor(0.67 * n_training)        # 2/3
  n_calibration     <- n_training - n_proper_training  # 1/3

  allocation <- rep(x     = c("proper training", "calibration"), 
                    times = c(n_proper_training, n_calibration)) %>%
  sample

@@

create_proper_training <- function(df) df %>% filter(allocation == "proper training")

create_calibration     <- function(df) df %>% filter(allocation == "calibration")
```

@@

  cv.glmnet(x = as.matrix(X), y = Y, type.measure = "mae") # lasso by default. Use mean absolute error

@@

  Y_hat <- predict.cv.glmnet(model, newx = as.matrix(X), s = "lambda.min") # use lambda that minimizes mae

  df %>% mutate(Y_hat = as.vector(Y_hat),
                error = abs(Y_hat - Y),
                mu    = log(error)) # take ln of absolute error

@@

  X  <- df %>% select(matches("GPA|Topic"))
  mu <- df %>% pull(mu)

  cv.glmnet(x = as.matrix(X), y = mu, type.measure = "mae")

@@

  mu_hat <- predict.cv.glmnet(model, newx = as.matrix(X), s = "lambda.min")

  df %>% mutate(mu_hat = as.vector(mu_hat))

@@

## Compute standardized non-conformity scores on callibration test
```{r}

compute_ncs              <- function(df) df %>% mutate(alpha = abs(Y - Y_hat))

compute_standardized_ncs <- function(df) df %>% mutate(sigma = exp(mu_hat),
                                                       alpha = abs(Y - Y_hat) / sigma )

compute_significance     <- function(df) df %>% mutate(significance = rank(-alpha) / nrow(.) ) # !!! double check!!!
```

## Evaluate error rate on test set

```{r}
compute_width <- function(df, significance_given){

  df <- df %>% filter(significance <= significance_given)

@@

  df %>% mutate(
    sigma       = exp(mu_hat),
    width       = ncs * sigma,
    border_low  = Y_hat - width,
    border_high = Y_hat + width
    )
}

@@

```{r, echo = FALSE}

significance <- c(0.05, seq(0.1, 0.9, 0.1))
results <- list()
n_list <- 1

for(course in course_selection){

@@

    # proper training, calibration and test sets
    Z_test            <- create_test(Z, fold_current = iteration)
    Z_training        <- create_training(Z, fold_current = iteration) %>% add_allocation
    Z_proper_training <- create_proper_training(Z_training)
    Z_calibration     <- create_calibration(Z_training)

    # fit target and error model to proper training set
    model_target      <- fit_target_model(Z_proper_training)
    Z_proper_training <- predict_grade(Z_proper_training, model_target)
    model_error       <- fit_error_model(Z_proper_training)

    # non-conformity scores on calibration
    Z_calibration <- Z_calibration %>% 
      predict_grade(model_target) %>%
      predict_error(model_error) %>%
      compute_standardized_ncs %>%
      compute_significance

    # error on test set
    Z_test <- Z_test %>% 
      predict_grade(model_target) %>%
      predict_error(model_error)

    for(signi in significance){ # loop through significance levels

      ncs <- compute_width(Z_calibration, signi)

      Z_test_augmented <- Z_test %>% 
        compute_interval(ncs = ncs) %>%
        compute_hit

      # save results
      results[[n_list]] <- Z_test_augmented %>%
        mutate(significance = signi,
               iteration    = iteration,
               course       = course) %>%
        select(course, iteration, significance, width, hit,
               Y, Y_hat, error, mu, mu_hat, sigma)

      n_list <- n_list + 1

    } # end loop significance

@@

```{r}
conformal_results <- bind_rows(results) %>%
    mutate(iteration    = as.factor(iteration),
           course       = as.factor(course))

save(conformal_results, course_table, file = "Output/conformal.RDATA")
```

Illustration of the effect of error_hat on interval's width (for significance of 99%):
```{r, fig.height=4}
Z_test %>% select(Y, Y_hat, error, error_hat, width, border_low, border_high, hit)
plot(Z_test$error, Z_test$error_hat)
```
